---
title: "Grid Edge Optimization"
author: "Brock Taute"
date: "4/30/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(CVXR)
library(tidyverse)
library(lubridate)

source('optimization_functions.R')
source('simulation_functions.R')

# 2018 electricity pricing and solar production data (hourly)
data <- read_csv('sample-data/data.csv')

```

# Background

This repo contains libraries to optimize an energy storage system's economic 
dispatch in grid "edge" scenarios.  It includes a convex problem formulation and
an interface to the CVXR package.  It's intended for prototyping and modeling 
different storage solutions and so emphasizes customizability over speed.  It 
includes functions to set up and simulate a testing environment where new
information is released periodically, such as with the Day Ahead pricing market. 
It also enables the analysis of sophisticated utility tariffs, including
multiple different demand charge rate structures based on seasons, day of the 
week, and monthly or daily billing periods.  Project configurations can range 
from standalone solar or storage projects to projects with solar, storage, and
a consumer load all on the same meter.

# Examples:

## 1. Remote Net Metering of a Solar + Storage Project

A remote net metering project fully utilizing the New York Value Stack Tariff
would have the goal of shifting solar production to take advantage of the 
variable Day Ahead electricity pricing.  Each day at 11AM, the following
day's pricing becomes available, which will give the economic dispatch engine 
an opportunity to update the optimal dispatch for the next 37 hours.  For this
example, a 5MW canopy solar system and a 5MW/15MWh energy storage system were
assumed to be connected with a grid injection limit of 5MW.  It was also assumed
that the storage system could only charge when offsetting solar production (it 
couldn't charge from the grid.)

Using Day Ahead electricity pricing and solar production from 2018, a full 
year's dispatch can be simulated at hourly intervals.

```{r example_1, results='hide'}

# Define the system

charge_efficiency <- .95
discharge_efficiency <- .95
storage_power <- 5000 # kW
storage_energy <- 15000 / discharge_efficiency # Nominal kWh
soc_start <- storage_energy
poi_limit <- 5000

# Defaults of function are for a NYSERDA, 1-Year simulation where
# Day Ahead pricing is released daily at 11AM
fs <- get_forecast_sequences()

rs <- simulate_remote_net_meter(fs$forecast_sequence, fs$forecast_lengths,
                             data$price, data$solar, storage_power,
                             storage_energy, poi_limit, charge_efficiency,
                             discharge_efficiency, soc_start)

df <- mutate(data,
             storage = rs$storage,
             soc = rs$soc,
             dispatch = solar - storage,
             value = price * dispatch)

```
```{r example_1_results, echo=FALSE}

cat(paste0('Predicted annual return:  $', round(sum(df$value), 0)), '\n',
    'Number of Storage Cycles per Year: ',
    round(-sum(df$storage[df$storage < 0]) / storage_energy, 1))

```

```{r example_1_average_plot, echo=FALSE}

# Apply a scaling factor to second y-axis to see pricing
price_scaling_factor <- storage_energy * .5

df %>%
  mutate(price = price * price_scaling_factor,
         month = month(datetime),
         hour = hour(datetime)) %>%
  group_by(month, hour) %>%
  summarise_all(mean) %>%
  gather('key', 'value', price, solar, storage, soc, dispatch) %>%
  ggplot(aes(x = hour, y = value, color = key)) +
  geom_path() +
  facet_wrap(~month) +
  scale_y_continuous(sec.axis = sec_axis(~./price_scaling_factor,
                                         name = 'Electricity Price ($/kWh)')) +
  scale_color_brewer(palette = 'Dark2') +
  scale_linetype_manual(values = c('solid', 'dotted', rep('solid', 3))) +
  labs(x = 'Datetime', y = 'kW', color = 'Key', linetype = 'Key',
       title = 'Average Hourly Dispatch Each Month') +
  theme_bw()

```

```{r example_1_timeseries_plot, echo=FALSE}

df %>%
  mutate(price = price * price_scaling_factor) %>%
  filter(datetime > ymd('2018/08/15'), datetime < ymd('2018/08/18')) %>%
  gather('key', 'value', solar, storage, soc, dispatch, price) %>%
  ggplot(aes(x = datetime, y = value, color = key)) +
  geom_path() +
  scale_y_continuous(sec.axis = sec_axis(~./price_scaling_factor,
                                         name = 'Electricity Price ($/kWh)')) +
  scale_color_brewer(palette = 'Dark2') +
  scale_linetype_manual(values = c('solid', 'dotted', rep('solid', 3))) +
  labs(x = 'Datetime', y = 'kW', color = 'Key', linetype = 'Key',
       title = 'Time Series Dispatch Mid-August') +
  theme_bw()

```

Looking at the dispatch plots, it is evident that the New York Value Stack 
Tariff should have the intended consequence of incentivizing solar production 
to be shifted toward the evenings.

## Example 2. Standalone Storage System with Multiple Demand Tariffs

For any system that consumes electricity from the grid, a utility bill must be 
considered in the dispatch algorithm.  For systems greater than a MW in size,
these tariffs can often be quite complex, with several different time-of-use 
structures.  To demonstrate how a tariff like this might be incoporated in the 
optimization, a tariff structure from ConEd is modeled for this example.  The 
tariff includes two different daily peak demand charges on top of the monthly
demand charge.  The model also appropriately accounts for taxes and fees tacked
onto the delivery charges.

Instead of modeling a daily dispatch for a year, this example allows the
optimization to incorporate data for the entire month of June at once.  This 
can be done to train dispatch algorithms.  For example, the optimal demand 
threshold values from this analysis can be used as set points for a daily
dispatch algorithm.

```{r example_2}

# Add in ConEd utility tariff to data

summer_months <- c(6, 7, 8, 9)
weekend <- c('Saturday', 'Sunday')
workweek <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')

full_data <- data %>%
  mutate(year = year(datetime),
         month = month(datetime),
         day = day(datetime),
         hour = hour(datetime),
         weekday = weekdays(datetime),
         # Constant monthly peak usage rate
         peakcost_monthly1 = 7.57,
         # Different Daily Peak rates based on day of week and season
         peakcost_daily1 = if_else(
           # If a weekday, 8am-10pm
           (hour >= 8) & (hour <= 22) & (weekday %in% workweek),
           if_else(month %in% summer_months, # And if in summer
                   1.0241, # then summer peak price
                   .7847, # else winter peak price
           ),
           # Else zero when not between 8am and 10pm (any time of year)
           0
         ),
         peakcost_daily2 = if_else(
           # If a summer weekday, 8am-6pm
           (month %in% summer_months) & (hour >= 8) & (hour <= 18) &
             (weekday %in% workweek),
           0.5298,
           0))

tax_kWh <- .024066
fees_kWh <- .005578

# Now prepare the data to model the optimal dispatch of a single month

model_month <- 6

# Filter out the unused months, then pivot the table wider twice, once for
# the monthly peaks and once for the daily peaks.  Each time use the abbreviated
# month name in the new column title.

test_data <- full_data %>%
  filter(month == model_month) %>%
  mutate(month = month.abb[month]) %>%
  pivot_wider(names_from = c(month, day), values_from = c(peakcost_daily1,
                                                          peakcost_daily2),
              values_fill = list(peakcost_daily1 = 0, peakcost_daily2 = 0)) %>%
  mutate(month = month.abb[month(datetime)]) %>%
  pivot_wider(names_from = 'month', values_from = peakcost_monthly1,
              values_fill = list(peakcost_monthly1 = 0),
              names_prefix = 'peakcost_monthly1_')

june_dispatch <- optimize_dispatch(storage_power, storage_energy, poi_limit,
                                   charge_efficiency, discharge_efficiency,
                                   soc_start, fees_kWh, tax_kWh,
                                   charge_solar_only = FALSE,
                                   price_kWh = test_data$price,
                                   peakcost_monthly1 =
                                     as.matrix(select(test_data,
                                                      matches('monthly1_Jun'))),
                                   peakcost_monthly2 = NULL,
                                   peakcost_daily1 = as.matrix(select(test_data,
                                                     matches('daily1_Jun'))),
                                   peakcost_daily2 = as.matrix(select(test_data,
                                                     matches('daily2_Jun'))))

# Calculate SOC using DC Side of storage dispatch

dc_storage <- june_dispatch

dc_storage[dc_storage > 0] <- dc_storage[dc_storage > 0] *
  charge_efficiency

dc_storage[dc_storage < 0] <- dc_storage[dc_storage < 0] /
  discharge_efficiency

soc <- cumsum(dc_storage) + soc_start

```

```{r example_2_plot, echo=FALSE}

# Visualize results

full_data %>%
  filter(month == 6) %>%
  mutate(storage = june_dispatch[,1],
         soc = soc,
         price = price * price_scaling_factor,
         peakcost_daily1 = peakcost_daily1 * price_scaling_factor,
         peakcost_daily2 = peakcost_daily2 * price_scaling_factor) %>%
  gather('key', 'value',
         soc, storage, price, peakcost_daily1, peakcost_daily2) %>%
  filter(datetime > ymd('2018/06/23'), datetime < ymd('2018/06/28')) %>%
  ggplot(aes(x = datetime, y = value, color = key)) +
  geom_path() +
  scale_y_continuous(sec.axis = sec_axis(~./price_scaling_factor,
                                         name = 'Electricity Price ($/kWh)')) +
  scale_color_brewer(palette = 'Dark2')+
  scale_linetype_manual(values = c('solid', 'dotted', rep('solid', 3))) +
  labs(x = 'Datetime', y = 'kW', color = 'Key', linetype = 'Key',
       title = 'Time Series Dispatch Mid-June') +
  theme_bw()

```

The plot shows two unique situations for the dispatch model.  Going into the 
weekend on June 24th, the storage system discharges completely and then uses the
two days without daily demand charges to fully recharge.  Then, on June 27th,
electricity prices sky rocket, and the storage system responds with a full
discharge again.  Between those events, the system maintains a higher state of
charge.  Under the optimal dispatch strategy with this tariff and incentive
structure, the energy storage system averages less than a daily cycle.

```{r example_2_cycles, echo=FALSE}

cat(paste0('Number of Storage Cycles in June: ',
           round(-sum(june_dispatch[june_dispatch < 0]) / storage_energy, 1)))

```